Location Classification on Tweets

### 2016-03-09 docs/paper.tex
rewritten data collection and filters

To be able to infer this information, we can try to use linguistic habits shared by the people who are in the same city. As humans, our spatial cognition deals with updating and managing the knowledge about our environment. Spatial cognition changes when we move between environments, but as we stay within the boundaries of a high level environment(i.e. a city), there may be some high level constants that doesn't change. If some of these constants are shared commonly between people, it may lead them to have similar linguistic habits, in this case, textual features in their tweets. If we group the geotagged tweets sent from a city, it may be possible to extract these features, if they exist.

Using similar features, previous research tries to determine the geolocation of a tweet using multi-class classification algorithms. These methods are trying to determine the most likely geolocation for a tweet from a list(e.g. cities in a country). It is sometimes hard to do that, because we need data from all cities. Also the algorithms and the implementations become more complicated. Thus, in cases where we just want to know if someone is from a city or not, this approach becomes unnecessarily heavy. In this research, we are simplifying this problem by using binary classification algorithms, which are trying to determine if a tweet belongs to a location(i.e. New York City) or not.

Revolving around our hypothesis, the research question we're trying to answer is: "Are there textual features in tweets, that are common for people in the same city? If there are, can they help us determine if a tweet is from that city, or not?"

@@

Using text mining methods and location tags in tweets, \cite{cheng2010you} redefined this problem. \cite{cheng2010you} used only the text data and introduced a method that identifies the words with strong local geo-scope. This study achieved up to 51\% accuracy to locate users within a 100 mile radius.

\cite{hecht2011tweets} introduced the $CALGARI$ classifying algorithm specific to this problem, focusing on country and state prediction. They got a 72\% accuracy for classifying 4 countries and 30\% accuracy on classifying 18 states.

\cite{eisenstein2010latent} created a geographic topic model to make use of the textual differences in topics based on regions, they got a median distance error of 494 kms.

\cite{wing2011simple} tried to solve this problem using geotagged Wikipedia articles and tweet texts, focusing on predicting the geolocation. Their work has obtained a median error of 479 km's for tweets.

\cite{mahmud2012tweet} introduced hierarchical classification methods combined with statistical methods. They have achieved 58\%, 66\% and 78\% accuracy on city, state and timezone levels, respectively. Also their findings show that users may be able to partially hide their locations, making their tweets untraceable. Same researchers revisited the problem in \cite{mahmud2014home}, trying to classify the home location of users, achieving similar results and also introducing movement and location prediction.

\cite{sadilek2012finding} approached this problem introducing the friend information, constructing a friend location graph. Their study was able to achieve 77\% to 84.3\% accuracy, relative to available number of friends.

\cite{graham2014world} studied the problems with the baselines accepted by recent studies and explained why location and language classification is not an easy problem and how some of the assumptions are incorrect, such as accepting the device location as user's location.

\section{Method}
\subsection{Data Collection}
Using Twitter's sample status streaming api\footnote{https://dev.twitter.com/streaming/reference/get/statuses/sample}, we have collected data. This api returns approximately 1\% of the public tweets from all around the world. We have filtered this data to reduce the problem space to only location classification. Our filter selects only the geotagged English tweets, sent within the United States. Running this collector for 1 month, we have collected a dataset consisting of 33112 individual tweets, each from a unique user. We have also collected a sample of 40 tweets from those users, if their timelines are public, to enrich the dataset.

Upon investigating this dataset, we have realised that there are a lot of low quality, automated tweets, coming from 886 different tweet sources(devices or applications that a tweet has been sent from). To remove these sources, we have created a whitelist of most used 20 sources that selects applications which real users use to tweet from. This reduced the amount of data by around 50\%, to 17040 individual geotagged tweets.

One problem with the locations is the granularity. A location can appear as a state(New York, US), a city(New York, NY) or a neighbourhood(Manhattan, NY). A selection of 8 top locations, not intersecting each other, is selected from the ones containing the most tweets.

>>>>

To be able to infer this information, we can try to use linguistic habits shared by the people who are within the same location. As humans, our spatial cognition deals with updating and managing the knowledge about our environment. Spatial cognition changes when we move between environments, but as we stay within the boundaries of a high level environment(i.e. state, city, borough), there may be some high-level constants that doesn't change. If some of these constants are shared commonly between people, it may lead them to have similar linguistic habits, in this case, textual features in their tweets. If we group the geotagged tweets sent within a high-level location, it may be possible to extract these features, if they exist.

Using similar features, previous research tries to determine the geolocation of a tweet using multi-class classification algorithms. These methods are trying to determine the most likely geolocation for a tweet from a list(e.g. cities in a country). It is sometimes hard to do that, because we need data from all cities. Also the algorithms and the implementations become more complicated. Thus, in cases where we just want to know if someone is within a state/city/borough or not, this approach becomes unnecessarily heavy. In this research, we are simplifying this problem by using binary classification algorithms, which are trying to determine if a tweet sent within the boundaries of a location(i.e. Manhattan, NY) or not.

Revolving around our hypothesis, the research question we're trying to answer is: "Are there textual features in tweets, that are common for people within the same location? If there are, can they help us determine if a tweet is sent within that high-level location, or not?"

@@

% Using text mining methods and location tags in tweets, \cite{cheng2010you} redefined this problem. \cite{cheng2010you} used only the text data and introduced a method that identifies the words with strong local geo-scope. This study achieved up to 51\% accuracy to locate users within a 100 mile radius.

% \cite{hecht2011tweets} introduced the $CALGARI$ classifying algorithm specific to this problem, focusing on country and state prediction. They got a 72\% accuracy for classifying 4 countries and 30\% accuracy on classifying 18 states.

% \cite{eisenstein2010latent} created a geographic topic model to make use of the textual differences in topics based on regions, they got a median distance error of 494 kms.

% \cite{wing2011simple} tried to solve this problem using geotagged Wikipedia articles and tweet texts, focusing on predicting the geolocation. Their work has obtained a median error of 479 km's for tweets.

% \cite{mahmud2012tweet} introduced hierarchical classification methods combined with statistical methods. They have achieved 58\%, 66\% and 78\% accuracy on city, state and timezone levels, respectively. Also their findings show that users may be able to partially hide their locations, making their tweets untraceable. Same researchers revisited the problem in \cite{mahmud2014home}, trying to classify the home location of users, achieving similar results and also introducing movement and location prediction.

% \cite{sadilek2012finding} approached this problem introducing the friend information, constructing a friend location graph. Their study was able to achieve 77\% to 84.3\% accuracy, relative to available number of friends.

% \cite{graham2014world} studied the problems with the baselines accepted by recent studies and explained why location and language classification is not an easy problem and how some of the assumptions are incorrect, such as accepting the device location as user's location.

\section{Method}
\subsection{Data Collection and Filtering}

Using Twitter's sample status streaming api\footnote{https://dev.twitter.com/streaming/reference/get/statuses/sample}, we have sampled the public tweets. This api samples approximately 1\% of all the public tweets around the world. To be able to find the tweets that are relevant to our task, we have applied some filters to this api.

First of all, we have applied a filter to ignore non-geotagged tweets. The high-level location information contained within each geotag will be used as the label of a tweet. These labels will be used to train our algorithms, and they will be compared with the predicted results to see how accurate our methods are.

Also, we are not interested with tweets in different languages. Having tweets from different languages would introduce a complicated variable and make our results hard to interpret. Therefore, we choose English as our language of interest, because we assumed that English is the most common language that people tweet, thus it would yield the biggest dataset(compared to other languages). Samples from twitter contain the language information, we have used this information to filter tweets which are not English.

Considering the language assumption, having tweets from non-English speaking countries would be unnecessary. Also, having tweets from different countries would introduce another variable. Assuming that United States would yield the biggest dataset(compared to other countries), we have limited the tweets we collected to United States.

Running the data collector for 1 month, we have collected a dataset consisting of 33112 tweets, geotagged within United States, tagged as English and each from a unique user.

Since each tweet is consisting of at most 140 characters, our dataset might not be enough to extract the textual features we're interested in. To enrich our database, we assumed that our users are not actively travelling. Therefore, we have collected past 40(at most) tweets of these users and attached them to the sampled tweet as history.

Assuming we had a dataset, big enough for the task, we examined a small sample of it. In this sample, we have seen many examples which were not composed by humans, but bots(e.g. "\textit{Police department activity on \#I95 SB at 3rd Avenue; Exit 3}"). Twitter also gives the information on which application is used to send a tweet. We have examined this information, called \textit{source}, to filter these tweets. Grouping and sorting them by the number of tweets, we were able to see which sources are most relevant for our study. Annotating the most relevant 20(e.g. \textit{Twitter for iPhone}, \textit{Facebook} and etc.) we have created the whitelist filter for tweet sources.

In our examination, we have also seen a variety of geotag names which represent the high-level locations we're interested in(e.g. \textit{Mississippi, USA}, \textit{Manhattan, NY} and etc.). When we checked the number of tweets per location, we have seen that there are 4898 unique locations in the whole dataset with different granularities(e.g. state, city, borough and etc). We have picked 8 of the most common locations(e.g. \textit{Manhattan, NY}, \textit{Los Angeles, CA}, \textit{Houston, TX} and etc.).

When we applied all these filters, we ended up having a dataset containing 8 locations, each of them having between 986 to 180 tweets.
